Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm:

### 2023-07-10 manuscript.Rmd
Added autobuild files

#delay
lure_delay_model1 <- buildmer(lure_recalled ~ Test.Time +intrusions +
                                       (intrusions | Participant.Private.ID) + 
                                       (Test.Time| answer), 
                                     data = lure_delay, 
                                     family = "binomial",
                                     buildmerControl = buildmerControl(direction='backward',args =list(control=glmerControl(optimizer="bobyqa"))))

@@

Our target sample size was 120 participants/group (i.e., 480 participants in total), defined as those who remained in the sample after applying the exclusion criteria outlined in section 9. This sample size gives us >=90% power to detect all the desired effects for our Research Questions (See Appendix D for a detailed power analysis).

@@

## Recruitment method. 
Following two previous sleep studies conducted via Prolific [@mak2023a; @mak2023b], we put a short survey on the platform to recruit a pool of participants (N = 2296). This survey is available in Appendix A and was hosted on Qualtrics. The first half of the survey asked for basic demographic information: gender identity, age, current country of residence, first language, ethnicity, highest education attainment, and history of developmental/sleep disorders (if any). The survey then provided a brief outline of the main study. It stated that if enrolled, participants would be randomly allocated to one of the four groups and that no preferences would be accommodated. Participants then indicated whether they would like to enrol in the main study. Of the 2296 respondents, 1940 expressed interest in taking part, who were then screened for their eligibility (see inclusion criteria below). Those who fitted our inclusion criteria were then randomly allocated to one of the four experimental groups. A private message was sent to each participant, notifying them of their group allocation. In the end, 534 participants completed both the study and test phases. These participants were reimbursed at a rate of ~£9.5/hr.

@@

We  made use of 20 DRM wordlists (see Table \@ref(tab:table2)), taken from Roediger et al. [-@roediger2001a]. Each list contained 8 semantically related words, and as per the standard DRM paradigm, they were arranged in a descending order of associative strength to the critical lures. A participant studied all 20 lists. We note that the original DRM lists by Roediger et al. [-@roediger2001a] were tailored for American participants, and two words (e.g., trash, Mississippi) were not immediately relatable to people in the UK. We, therefore, changed these words (e.g., trash → rubbish), as noted in Table \@ref(tab:table2).

@@

Immediately afterwards, participants in the AM/PM-control groups carried out the test phase. For those in the Delay groups, the test phase took place approximately 12 hours later. Here, both the Immediate and Delay participants rated their level of sleepiness on SSS and completed a short survey concerned with, for example, morningness/eveningness preference (rMEQ; [@adan1991a]) and sleep duration/quality the night before (see Appendix B for the full survey). This survey helped determine whether the four groups were matched in terms of time-of-day preference and whether data from a participant needed to be discarded as a result of meeting the exclusion criteria described in section 9. Finally, the test phase concluded with a 10-minute free recall task where participants recalled as many of the words as they could from the previously seen wordlists.

@@

There was a surprise attention check after the 4th, 9th, 13th, 18th lists, where participants saw an erroneous maths equation such as “3 + 3 = 11”. It was presented for 1 s, in the same font and style as the list words [@thomas2017a]. Immediately afterwards, participants were asked to report what 3 + 3 was according to what was just shown.

@@

4.	(Sleep and Wake groups only) they reported consuming any alcoholic drinks between study and test.

5.	(Sleep group only) they reported to have had fewer than 6 hours of overnight sleep prior to test or rated their sleep quality as poor or extremely poor.

6.	(Wake group only) they reported to have had a nap between study and test.

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance $\leq$ 2 (e.g., *cigerette → cigarette).\footnote{If a misspelt response has more than one nearest word, the response was considered as an intrusion.} Responses with added or dropped inflectional suffixes (i.e., -s, -ed, -ing, adjectival -er) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is pollution; if a participant recalled pollutions, the plural suffix was dropped; however, if a participant recalled pollutant, this was considered as an intrusion.

@@

A 2 $\times$ 2 Poisson regression revealed significant effects of Interval ($\beta$ = `r round(int_pois_sum$coefficients[2,1], digits = 3)`, SE = `r round(int_pois_sum$coefficients[2,2], digits = 3)`, z = `r round(int_pois_sum$coefficients[2,3], digits = 3)`, p < `r format.pval(int_pois_sum$coefficients[2,4], digits = 3, eps = .001)`) and Test Time ($\beta$ = `r round(int_pois_sum$coefficients[3,1], digits = 3)`, SE = `r round(int_pois_sum$coefficients[3,2], digits = 3)`, z = `r round(int_pois_sum$coefficients[3,3], digits = 3)`, p <  `r format.pval(int_pois_sum$coefficients[3,4], digits = 3, eps = .001)`), which were qualified by a significant interaction ($\beta$ = `r round(int_pois_sum$coefficients[4,1], digits = 3)`, SE =  `r round(int_pois_sum$coefficients[4,2], digits = 3)`, z = `r round(int_pois_sum$coefficients[4,3], digits = 3)`, p = `r round(int_pois_sum$coefficients[4,4], digits = 3)`). Given this, we tested the simple effects of Test Time within the Immediate and Delay groups using the emmeans package [@lenth2021a]. Within the Immediate groups, the evening participants (M = `r round(intrusion_CI[2,4], digits = 2)`; SD = `r round(intrusion_CI[2,5], digits = 2)`) produced more intrusions than the morning participants (M = `r round(intrusion_CI[1,4], digits = 2)`; SD = `r round(intrusion_CI[1,5], digits = 2)`) (z = `r round(int_sum$contrasts$z.ratio[1], digits = 3)`, p = `r format.pval(int_sum$contrasts$p.value[1], digits = 3, eps = .001)`). Likewise, in the Delay groups, the Wake participants, who completed free recall in the evening (M = `r round(intrusion_CI[4,4], digits = 2)`; SD = `r round(intrusion_CI[4,5], digits = 2)`), produced more intrusions than the Sleep participants, who completed recall in the morning (M = `r round(intrusion_CI[3,4], digits = 2)`; SD = `r round(intrusion_CI[3,5], digits = 2)`) (z = `r round(int_sum$contrasts$z.ratio[2], digits = 3)`, p = `r format.pval(int_sum$contrasts$p.value[2], digits = 3, eps = .001)`). Together, our data indicate that participants who attempted free recall in the evening (vs. morning) were more prone to intrusions, and this effect was greater in the Delay than in the Immediate groups. These unexpected findings prompted us to explore whether the number of total responses (i.e., studied + lures + intrusions) differed between morning and evening test time. Interestingly, this exploratory analysis (see section 13.1) showed no effect of Test Time. Together, these suggest that attempting free recall in the evening led to a selective increase in intrusions, but not necessarily a global increase in output bias. Finally, given that Test Time had a significant effect on intrusions, we followed our pre-registered analysis plan by adding However, neither Test Time nor their interaction is expected to be significant. If in the unlikely event that these are revealed to be significant, the number of intrusions as a numeric covariate in the 2 $\times$ 2 mixed-effects models below.

@@

A generalised linear mixed-effect model (GLMM) was fitted to the critical lure data on the item level (N of observations = `r nrow(lure_pp)`` participants $\times$ 20 critical lures).\footnote{The use of GLMM is a clear departure from prior ‘Sleep $\times$ DRM’ studies, the majority of which addressed the same research question using an independent t-test or ANOVA (e.g., Diekelmann et al., 2010; Monaghan et al., 2017; Payne et al., 2009). We explained in Appendix C why these statistical tests are usually not appropriate in the context of DRM recall and why GLMM are more advantageous.}  The dependent variable was binary: whether a critical lure was recalled or not (1 vs. 0). The fixed effects were the number of intrusions a participant produced, Interval (Immediate vs. Delay), Test Time (AM vs. PM), and an Interval by Test Time interaction. Interval and Test Time were coded using sum contrasts [@barr2019a]. The random-effect structure was determined by the “buildmer” package [@voeten2021a], which automatically found the maximal model that was capable of converging using backward elimination (with the “bobyqa” optimiser). This means that model selection started from the maximal model, as justified by the experimental design [@barr2013a]. The model we reported and based our interpretation on was the most maximal model that was capable of converging (see the upper half of Table \@ref(tab:table4) for the final random-effect structure and model output).

@@

The number of intrusions had a significant effect on lure recall, such that participants who produced more intrusions tended to recall more critical lures (z = `r round(lmout[4,3],digits=3)`, p < `r p_format(lmout[4,4],accuracy=0.001,digits=3,leading.zero=F)`). There were no main effects of Interval or Test Time (zs < 0.95, ps > .34), but there was a significant Interval by Test Time interaction (z = `r round(lmout[5,3],digits=3)`, p < `r p_format(lmout[5,4],accuracy=0.001,digits=2,leading.zero=F)`). Following our pre-registered analysis plan, we proceeded to test the simple effects of Test Time within the Immediate and Delay groups, using the emmeans package [@lenth2021a] in R. Among the Immediate groups, there was no significant difference in lure recall between the AM-control and PM-control participants ($\beta$ = `r emout1[6]`, SE = `r emout1[7]`, z = `r emout1[9]`, p = `r p_format(as.numeric(emout1[10]),leading.zero=F,digits=3)`). However, among the Delay groups, there was a significant difference ($\beta$ = `r emout2[7]`, SE = `r emout2[8]`, z = `r round(as.numeric(emout2[10]),digits=2)`, p = `r p_format(as.numeric(emout2[11]),leading.zero=F,digits=2)`) such that the Sleep participants (M = `r round(meanout1,digits=2)`, SD = `r round(sdout1,digits=2)`) reported more critical lures than the Wake participants (M = `r round(meanout2,digits=2)`, SD = `r round(sdout2,digits=2)`).

@@

We fitted a GLMM to the veridical recall dataset (N of observations = 488 participants $\times$ 160 studied words). The dependent variable was whether a studied word was recalled or not. The fixed effects were the the number of intrusions a participant produced, Interval (Immediate vs. Delay), Test Time (AM vs. PM), and an Interval by Test Time interaction. The coding scheme and computation procedure were the same as in the previous analysis. The model output and its random-effect structure are available in the lower half of Table \@ref(tab:table4). There were no significant effects of intrusions (z = `r table4bdata[3,4]`, p = `r table4bdata[3,5]`) or Test Time (z = `r table4bdata[5,4]`, p = `r table4bdata[5,5]`). However, there was a main effect of Interval (z = `r table4bdata[4,4]`, p `r p_format(smout[2,4],digits=3,accuracy=0.001,leading.zero=FALSE)`) such that participants in the Delay groups recalled significantly fewer studied words (M = `r round(mean(studied_delpp$Studied.Recalled),digits=2)`, SD = `r round(sd(studied_delpp$Studied.Recalled),digits=2)`) than those in the Immediate groups (M = `r round(mean(studied_impp$Studied.Recalled),digits=2)`, SD = `r round(sd(studied_impp$Studied.Recalled),digits=2)`), indicating time-dependent memory decay. Importantly, there was a significant Interval by Test Time interaction (z = `r table4bdata[6,4]`, p `r table4bdata[6,5]`), so we broke it down with the emmeans package as pre-registered. Within the Immediate groups, the evening participants (M = `r round(studied_CI[2,4],digits=2)`, SD = `r round(studied_CI[2,5],digits=2)`) recalled more studied words than the morning participants (M = `r round(studied_CI[1,4],digits=2)`, SD = `r round(studied_CI[1,5],digits=2)`), although this was not statistically significant ($\beta$ = `r emout3[6]`, SE = `r emout3[7]`, z = `r round(as.numeric(emout3[9]),digits=2)`, p = `r p_format(as.numeric(emout3[10]),leading.zero=F,digits=2)`). Within the Delay groups, there was a main effect of Test Time ($\beta$ = `r emout4[7]`, SE = `r emout4[8]`, z = `r round(as.numeric(emout4[10]),digits=3)`, p = `r p_format(as.numeric(emout4[11]),leading.zero=F,digits=2)`), such that the Sleep participants (M = `r round(studied_CI[3,4],digits=2)`, SD = `r round(studied_CI[3,5],digits=2)`) outperformed their Wake counterparts (M = `r round(studied_CI[4,4],digits=2)`, SD = `r round(studied_CI[4,5],digits=2)`). Together, these support the well-established findings that sleep is typically beneficial to the retention of newly encoded declarative memories.

@@

Bayes Factors were computed for (1) the Interval $\times$ Test Time interaction in the false and veridical mixed-effect models above, and for the simple effects of Test Time within the (2) Immediate and (3) Delay groups. Following the procedures in Gilbert et al. [@gilbert2018a], a Bayes Factor was computed using the Bayesian Information Criterion (BIC) approximation from two competing GLMMs. For instance, in computing the Bayes Factor for the Interval $\times$ Test Time interaction, two models were needed: An alternative model containing the full fixed-effects structure (Intrusions + Interval + Test Time + Interval $\times$ Test Time), and a null model lacking the interaction.\footnote{To obtain the Bayes Factor for the simple effects of Test Time, the alternative model will contain Test Time as the sole fixed effect while the null model will contain no fixed effects.} To estimate the Bayes Factor, we used the formula e $e{\Delta BIC_{10}/2}$, where $\Delta BIC_{10}$ is the BIC for the null model minus the BIC for the alternative model [@masson2011a; @lindelov2018a; @wagenmakers2007a]. This produces a Bayes Factor$_{10}$, which was interpreted with reference to Lee and Wagenmakers’ [-@lee2014a] heuristics. The current BIC approximation method has the advantage of being a straightforward solution for mixed-effects models; however, its usage remains controversial as it is known to favour the simpler model (i.e., the null hypothesis; [@lindelov2018a; @vandekerckhove2014a; @weakliem1999a]). Table \@ref(tab:table5) summarises the Bayes Factors derived from our mixed-effects models.

@@

Unlike the intrusion data which had an overall mean of `r round(mean(intrusions$intrusions),digits=1)` and a minimum of `r min(intrusions$intrusions)`, the number of total responses had a mean of `r round(mean(total$total),digits=1)` and a minimum of `r min(total$total)`, suggesting that it is better to consider total responses as continuous, as opposed to count, data. As such, we used a 2 $\times$ 2 between-participant ANOVA to test for the effects of Interval and Test Time on the number of total responses, which was log-transformed to give a more normal distribution. The ANOVA revealed a main effect of Interval [F(`r summary(total_anova)[[1]][[1]][1]`, `r summary(total_anova)[[1]][[1]][4]`) = `r round(summary(total_anova)[[1]][[4]][1],digits=2)`, p `r p_format(summary(total_anova)[[1]][[5]][1],digits=2,accuracy=0.001,leading.zero=F)`], such that participants in the Immediate groups (M = `r round(mean(total$total[which(total$Delay=='control')]),digits=1)`, SD = `r round(sd(total$total[which(total$Delay=='control')]),digits=1)`) gave more responses than those in the Delay groups (M = `r round(mean(total$total[which(total$Delay=='delay')]),digits=1)`, SD = `r round(sd(total$total[which(total$Delay=='delay')]),digits=1)`). This pattern is consistent with the confirmatory analysis on the studied list words. However, importantly, there was no significant effect of Test Time [F(`r summary(total_anova)[[1]][[1]][2]`, `r summary(total_anova)[[1]][[1]][4]`) = `r round(summary(total_anova)[[1]][[4]][2],digits=2)`, p = `r p_format(summary(total_anova)[[1]][[5]][2],digits=3,accuracy=0.001,leading.zero=F)`], and the Interval by Test Time interaction was also non-significant [F(`r summary(total_anova)[[1]][[1]][3]`, `r summary(total_anova)[[1]][[1]][4]`) = `r round(summary(total_anova)[[1]][[4]][3],digits=2)`, p = `r p_format(summary(total_anova)[[1]][[5]][3],digits=3,accuracy=0.001,leading.zero=F)`]. Together with the intrusion data, this exploratory analysis suggests that attempting free recall in the evening led to a selective increase in intrusions but not necessarily an increase in general response bias.

@@

Here, we asked whether recall probability of a lure (e.g., doctor) is predicted by the number of corresponding list items being recalled (e.g., nurse, hospital, sick), and if it does, whether it differs between the Sleep and Wake groups. These questions help shed light on the degree to which sleep increases lure recall via processes such as retrieval-induced generalisation\footnote{Retrieval-induced generalisation: Retrieval of one word cueing retrieval of a related word (e.g., Berens \& Bird, 2017)} or gist abstraction\footnote{Gist abstraction: Extraction of the central or essential meaning of learned information}, as these processes may predict a different degree of interdependence between lure and veridical recall. If sleep (vs. wake) promoted retrieval-induced generalisation, lure and veridical recall should become more strongly correlated with each other after sleep, because better veridical recall for a set of studied words may generalise to the corresponding critical lure (or vice versa). On the other hand, if sleep (vs. wake) promoted gist abstraction, lure recall may become less related to memories for the corresponding list items, because theories of gist abstraction such as iOtA may predict that sleep would selectively boost the overlapping gist memory (i.e., the lure) but not necessarily the specific studied words.

@@

As per the DRM literature, responses that were neither the studied list words nor the critical lures were classified as (non-critical) intrusions. For instance, our participants studied nurse, sick, lawyer, medicine, health, hospital, dentist, and physician, with doctor as the critical lure. Responses such as clinic and coconut would both be considered intrusions, but clearly, clinic is semantically more related to the studied list words than coconut. In other words, there is much diversity within the intrusion data. Here, we explored whether our four groups differed in terms of the semantic distance between intrusions and critical lures, as indexed by pre-trained semantic spaces (ukWaC; [@baroni2009a]) derived from word2vec [@guenther2015a]. We reasoned that since lure recall was greater in our Sleep (vs. Wake) participants, the intrusions produced by these participants could be more related to the lures in semantic space (e.g., @mak2023a). To test this, we computed the cosine similarities between each intrusion and each of the 20 critical lures (see Table \@ref(tab:table8) for an illustration). The intrusion-lure pair with the highest cosine similarity (i.e., the nearest neighbour) was used for this analysis.

>>>>

#delay
lure_delay_model1 <- buildmer(lure_recalled ~ Test.Time +intrusions + (intrusions | Participant.Private.ID) + (Test.Time| answer), data = lure_delay, family = "binomial", buildmerControl = buildmerControl(direction='backward',args = list(control=glmerControl(optimizer="bobyqa"))))

@@

Our target sample size was 120 participants/group (i.e., 480 participants in total), defined as those who remained in the sample after applying the exclusion criteria outlined in section 9. This sample size gives us >=90% power to detect all the desired effects for our Research Questions (See **Appendix D** for a detailed power analysis).

@@

## Recruitment method. 
Following two previous sleep studies conducted via Prolific [@mak2023a; @mak2023b], we put a short survey on the platform to recruit a pool of participants (N = 2296). This survey is available in **Appendix A** and was hosted on Qualtrics. The first half of the survey asked for basic demographic information: gender identity, age, current country of residence, first language, ethnicity, highest education attainment, and history of developmental/sleep disorders (if any). The survey then provided a brief outline of the main study. It stated that if enrolled, participants would be randomly allocated to one of the four groups and that no preferences would be accommodated. Participants then indicated whether they would like to enrol in the main study. Of the 2296 respondents, 1940 expressed interest in taking part, who were then screened for their eligibility (see inclusion criteria below). Those who fitted our inclusion criteria were then randomly allocated to one of the four experimental groups. A private message was sent to each participant, notifying them of their group allocation. In the end, 534 participants completed both the study and test phases. These participants were reimbursed at a rate of ~£9.5/hr.

@@

We  made use of 20 DRM wordlists (see Table \@ref(tab:table2)), taken from Roediger et al. [-@roediger2001a]. Each list contained 8 semantically related words, and as per the standard DRM paradigm, they were arranged in a descending order of associative strength to the critical lures. A participant studied all 20 lists. We note that the original DRM lists by Roediger et al. [-@roediger2001a] were tailored for American participants, and two words (e.g., *trash*, *Mississippi*) were not immediately relatable to people in the UK. We, therefore, changed these words (e.g., *trash* → *rubbish*), as noted in Table \@ref(tab:table2).

@@

Immediately afterwards, participants in the AM/PM-control groups carried out the test phase. For those in the Delay groups, the test phase took place approximately 12 hours later. Here, both the Immediate and Delay participants rated their level of sleepiness on SSS and completed a short survey concerned with, for example, morningness/eveningness preference (rMEQ; [@adan1991a]) and sleep duration/quality the night before (see **Appendix B** for the full survey). This survey helped determine whether the four groups were matched in terms of time-of-day preference and whether data from a participant needed to be discarded as a result of meeting the exclusion criteria described in section 9. Finally, the test phase concluded with a 10-minute free recall task where participants recalled as many of the words as they could from the previously seen wordlists.

@@

There was a surprise attention check after the 4$^{th}$, 9$^{th}$, 13$^{th}$, 18$^{th}$ lists, where participants saw an erroneous maths equation such as “3 + 3 = 11”. It was presented for 1 s, in the same font and style as the list words [@thomas2017a]. Immediately afterwards, participants were asked to report what 3 + 3 was according to what was just shown.

@@

4.	(*Sleep and Wake groups only*) they reported consuming any alcoholic drinks between study and test.

5.	(*Sleep group only*) they reported to have had fewer than 6 hours of overnight sleep prior to test or rated their sleep quality as poor or extremely poor.

6.	(*Wake group only*) they reported to have had a nap between study and test.

@@

The free recall data were pre-processed. The first step was to remove any duplicate responses. The second was to correct all obvious spelling and typing errors to the nearest English words, defined as Levenstein distance $\leq$ 2 (e.g., **cigerette* → *cigarette*).\footnote{If a misspelt response has more than one nearest word, the response was considered as an intrusion.} Responses with added or dropped inflectional suffixes (i.e., *-s*, *-ed*, *-ing*, adjectival *-er*) were corrected. Responses with derivational changes were considered as intrusions. For instance, one of the studied words is *pollution*; if a participant recalled *pollutions*, the plural suffix was dropped; however, if a participant recalled *pollutant*, this was considered as an intrusion.

@@

A 2 $\times$ 2 Poisson regression revealed significant effects of Interval ($\beta$ = `r round(int_pois_sum$coefficients[2,1], digits = 3)`, SE = `r round(int_pois_sum$coefficients[2,2], digits = 3)`, z = `r round(int_pois_sum$coefficients[2,3], digits = 3)`, p < `r format.pval(int_pois_sum$coefficients[2,4], digits = 3, eps = .001)`) and Test Time ($\beta$ = `r round(int_pois_sum$coefficients[3,1], digits = 3)`, SE = `r round(int_pois_sum$coefficients[3,2], digits = 3)`, z = `r round(int_pois_sum$coefficients[3,3], digits = 3)`, p <  `r format.pval(int_pois_sum$coefficients[3,4], digits = 3, eps = .001)`), which were qualified by a significant interaction ($\beta$ = `r round(int_pois_sum$coefficients[4,1], digits = 3)`, SE =  `r round(int_pois_sum$coefficients[4,2], digits = 3)`, z = `r round(int_pois_sum$coefficients[4,3], digits = 3)`, p = `r round(int_pois_sum$coefficients[4,4], digits = 3)`). Given this, we tested the simple effects of Test Time within the Immediate and Delay groups using the emmeans package [@lenth2021a]. Within the Immediate groups, the evening participants (M = `r round(intrusion_CI[2,4], digits = 2)`; SD = `r round(intrusion_CI[2,5], digits = 2)`) produced more intrusions than the morning participants (M = `r round(intrusion_CI[1,4], digits = 2)`; SD = `r round(intrusion_CI[1,5], digits = 2)`) (z = `r round(int_sum$contrasts$z.ratio[1], digits = 3)`, p = `r format.pval(int_sum$contrasts$p.value[1], digits = 3, eps = .001)`). Likewise, in the Delay groups, the Wake participants, who completed free recall in the evening (M = `r round(intrusion_CI[4,4], digits = 2)`; SD = `r round(intrusion_CI[4,5], digits = 2)`), produced more intrusions than the Sleep participants, who completed recall in the morning (M = `r round(intrusion_CI[3,4], digits = 2)`; SD = `r round(intrusion_CI[3,5], digits = 2)`) (z = `r round(int_sum$contrasts$z.ratio[2], digits = 3)`, p = `r format.pval(int_sum$contrasts$p.value[2], digits = 3, eps = .001)`). Together, our data indicate that participants who attempted free recall in the evening (vs. morning) were more prone to intrusions, and this effect was greater in the Delay than in the Immediate groups. These unexpected findings prompted us to explore whether *the number of total responses* (i.e., studied + lures + intrusions) differed between morning and evening test time. Interestingly, this exploratory analysis (see section 13.1) showed no effect of Test Time. Together, these suggest that attempting free recall in the evening led to a selective increase in intrusions, but not necessarily a global increase in output bias. Finally, given that Test Time had a significant effect on intrusions, we followed our pre-registered analysis plan by adding the number of intrusions as a numeric covariate in the 2 $\times$ 2 mixed-effects models below.

@@

A generalised linear mixed-effect model (GLMM) was fitted to the critical lure data on the item level (N of observations = `r nrow(lure_pp)` participants $\times$ 20 critical lures).\footnote{The use of GLMM is a clear departure from prior ‘Sleep $\times$ DRM’ studies, the majority of which addressed the same research question using an independent t-test or ANOVA (e.g., Diekelmann et al., 2010; Monaghan et al., 2017; Payne et al., 2009). We explained in Appendix C why these statistical tests are usually not appropriate in the context of DRM recall and why GLMM are more advantageous.}  The dependent variable was binary: whether a critical lure was recalled or not (1 vs. 0). The fixed effects were the number of intrusions a participant produced, Interval (Immediate vs. Delay), Test Time (AM vs. PM), and an Interval by Test Time interaction. Interval and Test Time were coded using sum contrasts [@barr2019a]. The random-effect structure was determined by the “buildmer” package [@voeten2021a], which automatically found the maximal model that was capable of converging using backward elimination (with the “bobyqa” optimiser). This means that model selection started from the maximal model, as justified by the experimental design [@barr2013a]. The model we reported and based our interpretation on was the most maximal model that was capable of converging (see the upper half of Table \@ref(tab:table4) for the final random-effect structure and model output).

@@

The number of intrusions had a significant effect on lure recall, such that participants who produced more intrusions tended to recall more critical lures (z = `r round(lmout[4,3],digits=3)`, p < `r p_format(lmout[4,4],accuracy=0.001,digits=3,leading.zero=F)`). There were no main effects of Interval or Test Time (zs < 0.95, ps > .34), but there was a significant Interval by Test Time interaction (z = `r round(lmout[5,3],digits=3)`, p = `r p_format(lmout[5,4],accuracy=0.001,digits=2,leading.zero=F)`). Following our pre-registered analysis plan, we proceeded to test the simple effects of Test Time within the Immediate and Delay groups, using the emmeans package [@lenth2021a] in R. Among the Immediate groups, there was no significant difference in lure recall between the AM-control and PM-control participants ($\beta$ = `r emout1[6]`, SE = `r emout1[7]`, z = `r emout1[9]`, p = `r p_format(as.numeric(emout1[10]),leading.zero=F,digits=3)`). However, among the Delay groups, there was a significant difference ($\beta$ = `r emout2[7]`, SE = `r emout2[8]`, z = `r round(as.numeric(emout2[10]),digits=2)`, p = `r p_format(as.numeric(emout2[11]),leading.zero=F,digits=2)`) such that the Sleep participants (M = `r round(meanout1,digits=2)`, SD = `r round(sdout1,digits=2)`) reported more critical lures than the Wake participants (M = `r round(meanout2,digits=2)`, SD = `r round(sdout2,digits=2)`).

@@

We fitted a GLMM to the veridical recall dataset (N of observations = 488 participants $\times$ 160 studied words). The dependent variable was whether a studied word was recalled or not. The fixed effects were the the number of intrusions a participant produced, Interval (Immediate vs. Delay), Test Time (AM vs. PM), and an Interval by Test Time interaction. The coding scheme and computation procedure were the same as in the previous analysis. The model output and its random-effect structure are available in the lower half of Table \@ref(tab:table4). There were no significant effects of intrusions (z = `r table4bdata[3,4]`, p = `r table4bdata[3,5]`) or Test Time (z = `r table4bdata[5,4]`, p = `r table4bdata[5,5]`). However, there was a main effect of Interval (z = `r table4bdata[4,4]`, p `r p_format(smout[2,4],digits=3,accuracy=0.001,leading.zero=FALSE)`) such that participants in the Delay groups recalled significantly fewer studied words (M = `r round(mean(studied_delpp$Studied.Recalled),digits=2)`, SD = `r round(sd(studied_delpp$Studied.Recalled),digits=2)`) than those in the Immediate groups (M = `r round(mean(studied_impp$Studied.Recalled),digits=2)`, SD = `r round(sd(studied_impp$Studied.Recalled),digits=2)`), indicating time-dependent memory decay. Importantly, there was a significant Interval by Test Time interaction (z = `r table4bdata[6,4]`, p `r p_format(smout[5,4],digits=2,accuracy=0.001,leading.zero=FALSE)`), so we broke it down with the emmeans package as pre-registered. Within the Immediate groups, the evening participants (M = `r round(studied_CI[2,4],digits=2)`, SD = `r round(studied_CI[2,5],digits=2)`) recalled more studied words than the morning participants (M = `r round(studied_CI[1,4],digits=2)`, SD = `r round(studied_CI[1,5],digits=2)`), although this was not statistically significant ($\beta$ = `r emout3[6]`, SE = `r emout3[7]`, z = `r round(as.numeric(emout3[9]),digits=2)`, p = `r p_format(as.numeric(emout3[10]),leading.zero=F,digits=2)`). Within the Delay groups, there was a main effect of Test Time ($\beta$ = `r emout4[7]`, SE = `r emout4[8]`, z = `r round(as.numeric(emout4[10]),digits=3)`, p = `r p_format(as.numeric(emout4[11]),leading.zero=F,digits=1)`), such that the Sleep participants (M = `r round(studied_CI[3,4],digits=2)`, SD = `r round(studied_CI[3,5],digits=2)`) outperformed their Wake counterparts (M = `r round(studied_CI[4,4],digits=2)`, SD = `r round(studied_CI[4,5],digits=2)`). Together, these support the well-established findings that sleep is typically beneficial to the retention of newly encoded declarative memories.

@@

Bayes Factors were computed for (1) the Interval $\times$ Test Time interaction in the false and veridical mixed-effect models above, and for the simple effects of Test Time within the (2) Immediate and (3) Delay groups. Following the procedures in Gilbert et al. [@gilbert2018a], a Bayes Factor was computed using the Bayesian Information Criterion (BIC) approximation from two competing GLMMs. For instance, in computing the Bayes Factor for the Interval $\times$ Test Time interaction, two models were needed: An alternative model containing the full fixed-effects structure (Intrusions + Interval + Test Time + Interval $\times$ Test Time), and a null model lacking the interaction.\footnote{To obtain the Bayes Factor for the simple effects of Test Time, the alternative model will contain Test Time as the sole fixed effect while the null model will contain no fixed effects.} To estimate the Bayes Factor, we used the formula $e^{\Delta BIC_{10}/2}$, where $\Delta BIC_{10}$ is the BIC for the null model minus the BIC for the alternative model [@masson2011a; @lindelov2018a; @wagenmakers2007a]. This produces a Bayes Factor$_{10}$, which was interpreted with reference to Lee and Wagenmakers’ [-@lee2014a] heuristics. The current BIC approximation method has the advantage of being a straightforward solution for mixed-effects models; however, its usage remains controversial as it is known to favour the simpler model (i.e., the null hypothesis; [@lindelov2018a; @vandekerckhove2014a; @weakliem1999a]). Table \@ref(tab:table5) summarises the Bayes Factors derived from our mixed-effects models.

@@

Unlike the intrusion data which had an overall mean of `r round(mean(intrusions$intrusions),digits=1)` and a minimum of `r min(intrusions$intrusions)`, the number of total responses had a mean of `r round(mean(total$total),digits=1)` and a minimum of `r min(total$total)`, suggesting that it is better to consider total responses as continuous, as opposed to count, data. As such, we used a 2 $\times$ 2 between-participant ANOVA to test for the effects of Interval and Test Time on the number of total responses, which was log-transformed to give a more normal distribution. The ANOVA revealed a main effect of Interval [F(`r summary(total_anova)[[1]][[1]][1]`, `r summary(total_anova)[[1]][[1]][4]`) = `r round(summary(total_anova)[[1]][[4]][1],digits=2)`, p `r p_format(summary(total_anova)[[1]][[5]][1],digits=2,accuracy=0.001,leading.zero=F)`], such that participants in the Immediate groups (M = `r round(mean(total$total[which(total$Delay=='Immediate/Control')]),digits=1)`, SD = `r round(sd(total$total[which(total$Delay=='Immediate/Control')]),digits=1)`) gave more responses than those in the Delay groups (M = `r round(mean(total$total[which(total$Delay=='12-hour delay')]),digits=1)`, SD = `r round(sd(total$total[which(total$Delay=='12-hour delay')]),digits=1)`). This pattern is consistent with the confirmatory analysis on the studied list words. However, importantly, there was no significant effect of Test Time [F(`r summary(total_anova)[[1]][[1]][2]`, `r summary(total_anova)[[1]][[1]][4]`) = `r round(summary(total_anova)[[1]][[4]][2],digits=2)`, p = `r p_format(summary(total_anova)[[1]][[5]][2],digits=3,accuracy=0.001,leading.zero=F)`], and the Interval by Test Time interaction was also non-significant [F(`r summary(total_anova)[[1]][[1]][3]`, `r summary(total_anova)[[1]][[1]][4]`) = `r round(summary(total_anova)[[1]][[4]][3],digits=2)`, p = `r p_format(summary(total_anova)[[1]][[5]][3],digits=3,accuracy=0.001,leading.zero=F)`]. Together with the intrusion data, this exploratory analysis suggests that attempting free recall in the evening led to a selective increase in intrusions but not necessarily an increase in general response bias.

@@

Here, we asked whether recall probability of a lure (e.g., *doctor*) is predicted by the number of corresponding list items being recalled (e.g., *nurse*, *hospital*, *sick*), and if it does, whether it differs between the Sleep and Wake groups. These questions help shed light on the degree to which sleep increases lure recall via processes such as retrieval-induced generalisation\footnote{Retrieval-induced generalisation: Retrieval of one word cueing retrieval of a related word (e.g., Berens \& Bird, 2017)} or gist abstraction\footnote{Gist abstraction: Extraction of the central or essential meaning of learned information}, as these processes may predict a different degree of interdependence between lure and veridical recall. If sleep (vs. wake) promoted retrieval-induced generalisation, lure and veridical recall should become more strongly correlated with each other after sleep, because better veridical recall for a set of studied words may generalise to the corresponding critical lure (or vice versa). On the other hand, if sleep (vs. wake) promoted gist abstraction, lure recall may become less related to memories for the corresponding list items, because theories of gist abstraction such as iOtA may predict that sleep would selectively boost the overlapping gist memory (i.e., the lure) but not necessarily the specific studied words.

@@

As per the DRM literature, responses that were neither the studied list words nor the critical lures were classified as (non-critical) intrusions. For instance, our participants studied *nurse*, *sick*, *lawyer*, *medicine*, *health*, *hospital*, *dentist*, and *physician*, with *doctor* as the critical lure. Responses such as *clinic* and *coconut* would both be considered intrusions, but clearly, *clinic* is semantically more related to the studied list words than *coconut*. In other words, there is much diversity within the intrusion data. Here, we explored whether our four groups differed in terms of the semantic distance between intrusions and critical lures, as indexed by pre-trained semantic spaces (ukWaC; [@baroni2009a]) derived from word2vec [@guenther2015a]. We reasoned that since lure recall was greater in our Sleep (vs. Wake) participants, the intrusions produced by these participants could be more related to the lures in semantic space (e.g., @mak2023a). To test this, we computed the cosine similarities between each intrusion and each of the 20 critical lures (see Table \@ref(tab:table8) for an illustration). The intrusion-lure pair with the highest cosine similarity (i.e., the nearest neighbour) was used for this analysis.
